Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-11-01 response_to_reviewers.Rmd

> Table 2 seems unnecessary, as the content is entirely summarized in the text.

We have removed Table 2.

> After the comprehensive views in Sections 2 and 3, Section 4 seems sparse on details. Much detail is given on the Bradley-Terry model, but only one example is given of a possible usage. Has this been used in other sports? Has it changed how we think about those sports or team strategies? An explanation of Elo rating should be given; I for one have no knowledge of how to compute it.

We thank the reviewer for this excellent suggestion. We have added a subsection on Elo ratings, with a mathematical definition followed by a calculation example. This example uses the `chessR` package, which ties into and strengthens our discussion of tools in the next chapter.

@@

> 11, 36: Table 3, why order this table alphabetically, why not by ability level?

We have ordered the rows of Table 3 by descending order of ability.

> 11, 50: Given how much fivethirtyeight.com uses Elo ratings for all kinds of sports, it is probably worth mentioning their work as examples.

We have added the sentence: "Elo ratings are used extensively for rankings of teams in numerous sports by the data journalists at FiveThirtyEight.com."

### 2022-11-01 sports_analytics.Rmd
close #36

To illustrate, we consider a chess game played on June 1, 2022 on [Chess.com](https://www.chess.com/) by one of the authors, with data obtained from the \pkg{chessR} package [@R-chessR]. 
Prior to the game, the author was rated `r greg[6, "BlackElo"]`, whereas `r greg[6, "WhiteElo"]` was the opponent's rating.
Since both ratings are below 2400, we apply a development coefficient of $K=20$ to this example.
The probability of author ($a$) defeating their opponent ($b$) is
<!-- don't know how to include `r` inside latex -->
$$
\Pi_{ab} = \frac{1}{1 + 20^{(1683 - 1732)/400}} = 0.591 \,.
$$
The playing author finished with a victory, which is associated with a score of $S_a = 1$.
The post-game Elo rating for this author is
$$
R^*_a = 1732 + 20(1 - 0.591) = 1740 \,.
$$

@@

## Case study in how tools fit together: chess

>>>>

To illustrate, we consider a chess game played on June 1, 2022 on [Chess.com](https://www.chess.com/) by one of the authors, with data obtained from the \pkg{chessR} package [@R-chessR] (see Section \ref{sec:chess}). 
Prior to the game, the author was rated `r greg[6, "BlackElo"]`, whereas his opponent was rated `r greg[6, "WhiteElo"]`.
Since both ratings are below 2400, we apply a development coefficient of $K=20$ to this example.
The probability of the author ($a$) defeating their opponent ($b$) was
<!-- don't know how to include `r` inside latex -->
$$
\Pi_{ab} = \frac{1}{1 + 20^{(1683 - 1732)/400}} = 0.591 \,.
$$
The author won the match: that outcome is associated with a score of $S_a = 1$.
The post-game Elo rating for the author is thus
$$
R^*_a = 1732 + 20(1 - 0.591) = 1740 \,.
$$

@@

## Case study in how tools fit together: chess {#sec:chess}
